---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
path=getwd()
zip_file=file.path(getwd(), "activity.zip")
unzip(zipfile = zip_file)
data=read.csv("activity.csv")
str(data)
data$date=as.Date(as.character(data$date), "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
```{r}
library(plyr)
library(ggplot2)
mn=ddply(data, "date", summarise, steps = sum(steps, na.rm = TRUE))
ggplot(mn, aes(steps)) + 
        geom_histogram(bins = 6) +
        ggtitle("Total number of steps taken each day") + 
        xlab("Number of steps") + ylab("Frequency")
```

```{r meanmed, echo=FALSE}
men=round(mean(mn[,2]),2)
med=median(mn[,2])
```

Mean and Median of the Total number of steps taken per day are `r men` and `r med` respectively.

## What is the average daily activity pattern?
```{r}
minst= as.data.frame(tapply(data$steps, data$interval, mean, na.rm = T))
int=subset(data, data$date=="2012-10-01")
int=int[,3]
tot=cbind(minst, int)
plot(tot[,2], tot[,1], type='l', xlab="5 minute interval", ylab="average number of steps", main="time series plot")
```

```{r av, echo=FALSE}
ma=row(tot)[tot==max(tot[,1])]
maks=max(tot[,1])
```

5-minute interval is `r ma` with the maximum number of `r maks` steps

## Imputing missing values

```{r na}
d1=sum(is.na(data$steps==TRUE))
d2=sum(is.na(data$date==TRUE))
d3=sum(is.na(data$interval==TRUE))
imp=sum(apply(data, 1, anyNA))
```

NA values exists only in steps data where total amount is `r d1` so we have to replace NA values only in steps data. Date and interval data has no NA values: `r d2` and `r d3` respectively. Thus, the total value of NA values is `r imp`. 

As median of steps data is 0, I will use mean of the steps data to replace the NA values in the steps dataset. 

```{r}
data1=data
for (i in 1:length(data[,1])){
        if(is.na(data1[i,1])==TRUE){
        data1[i,1]=mean(data[,1], na.rm = TRUE)
        }
}
sum(is.na(data1))
```

Histogram of total number of steps taken each day:
```{r}
mn1=ddply(data1, "date", summarise, steps = sum(steps))
ggplot(mn1, aes(steps)) + 
        geom_histogram(bins = 5) +
        ggtitle("Total number of steps taken each day") + 
        xlab("Number of steps") + ylab("Frequency")
```

```{r meanmed1, echo=FALSE}
options(scipen=9999)
men1=round(mean(mn1[,2]),2)
med1=round(median(mn1[,2]),2)
```

Mean and Median of the Total number of steps taken per day are `r men1` and `r med1` respectively.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
dd=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
wkd1=weekdays(data[,2])
wkd=ifelse(wkd1 %in% dd, "weekday", "weekend")
wkd=as.factor(wkd)
data1=cbind(data1, wkd)
colnames(data1)[4]="day"
```

Weekdays and weekend data was added to the previously created dataset and below is the first 6 lines of new created dataset:

```{r fin, echo = FALSE}
head(data1)
``` 

Final plot of the average steps by the weekdays and weekends: 

``` {r}
final1=ddply(data1, .(interval, day), summarise, steps = mean(steps))
qplot(x=interval, y=steps, data=final1, facets=.~day, geom="line", xlab="5 minute interval", ylab="Averarge number of steps", main="Average number of steps by the weekdays and weekends")
```
